---
title: "qPCR Filtered Samples Host-DNA Depletion Analysis"
author: "Your Name"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: flatly
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 12,
  fig.height = 10
)
```

```{r load libraries}
# Load required libraries
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(rstatix)
library(readxl)
library(patchwork)
library(writexl)
```

# Overview

This analysis examines host-DNA depletion efficacy across filtered samples for three organisms:
- **Chicken** DNA (host DNA being depleted)
- **Campylobacter** DNA (target pathogen)
- **Salmonella** DNA (target pathogen)

Two filtration methods were tested:
- **Cent + Filt**: Centrifugation followed by filtration
- **Filt**: Filtration only

---

# Chicken DNA Analysis

## Import and Clean Data

```{r chicken-data}
# Set detection limits
detection_limit_ct <- 40
detection_limit_conc <- 0.00001

# Read chicken data
chicken_raw <- read_excel('/Volumes/Alice/1_METHOD_DEVELOPMENT/Host-DNA_depletion/Filtered_samples_qPCR/qPCR_filtered-host-dep_Chicken.xlsx')

# Clean data
chicken_clean <- chicken_raw %>%
  filter(!is.na(Name), !str_detect(Name, "Blank")) %>%
  mutate(
    # Extract treatment
    Treatment = if_else(str_detect(Name, "No Host"), "Before Depletion", "After Depletion"),
    Treatment = factor(Treatment, levels = c("Before Depletion", "After Depletion")),
    # Extract sample type
    Sample_Type = case_when(
      str_detect(Name, "Cent \\+ Filt") ~ "Centrifuge + Filter",
      str_detect(Name, "^Filt") ~ "Filter Only",
      TRUE ~ "Unknown"
    ),
    # Extract replicate ID
    Replicate_ID = as.numeric(str_extract(Name, "\\d+")),
    # Non-detects
    Below_Detection = is.na(Ct) | is.na(`Calc Conc (ng/ul)`),
    # For plotting
    Ct_plot = ifelse(is.na(Ct), detection_limit_ct, Ct),
    Conc_plot = ifelse(is.na(`Calc Conc (ng/ul)`), detection_limit_conc, `Calc Conc (ng/ul)`)
  ) %>%
  select(Name, Treatment, Sample_Type, Replicate_ID, Ct, Ct_plot, 
         Concentration = `Calc Conc (ng/ul)`, Conc_plot, Below_Detection)

# Summary statistics
chicken_summary <- chicken_clean %>%
  group_by(Sample_Type, Treatment, Replicate_ID) %>%
  summarise(
    n_measurements = n(),
    n_detected = sum(!Below_Detection),
    n_non_detect = sum(Below_Detection),
    All_Non_Detect = all(Below_Detection),
    Mean_Ct = ifelse(all(is.na(Ct)), NA_real_, mean(Ct, na.rm = TRUE)),
    SE_Ct = ifelse(all(is.na(Ct)), NA_real_, sd(Ct, na.rm = TRUE) / sqrt(sum(!is.na(Ct)))),
    Mean_Conc = ifelse(all(is.na(Concentration)), NA_real_, mean(Concentration, na.rm = TRUE)),
    SE_Conc = ifelse(all(is.na(Concentration)), NA_real_, sd(Concentration, na.rm = TRUE) / sqrt(sum(!is.na(Concentration)))),
    .groups = "drop"
  )

chicken_summary %>%
  select(Sample_Type, Treatment, Replicate_ID, n_detected, n_non_detect, Mean_Ct, Mean_Conc) %>%
  knitr::kable(digits = 4, caption = "Chicken DNA: Summary by Sample Type and Replicate")
```

## Combined Panel Plot

```{r chicken-plot}
# Create long format
chicken_long <- chicken_clean %>%
  select(Sample_Type, Treatment, Replicate_ID, Below_Detection, Ct_plot, Conc_plot) %>%
  pivot_longer(
    cols = c(Ct_plot, Conc_plot),
    names_to = "Measurement",
    values_to = "Value"
  ) %>%
  mutate(
    Measurement = factor(Measurement,
                        levels = c("Ct_plot", "Conc_plot"),
                        labels = c("Ct Value", 
                                  "DNA Concentration (ng/µL)"))
  )

# Color palette
color_palette_chicken <- c("1" = "#88D498", "2" = "#A846A0", "3" = "#98C1D9")

# Create plot
chicken_panel <- ggplot(chicken_long,
                        aes(x = Treatment, y = Value,
                            color = as.factor(Replicate_ID),
                            group = as.factor(Replicate_ID))) +
  stat_summary(fun = mean, geom = "line", linewidth = 1.2, na.rm = TRUE,
               position = position_dodge(width = 0.3)) +
  stat_summary(fun = mean, geom = "point", size = 3.5, na.rm = TRUE,
               position = position_dodge(width = 0.3)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", 
               width = 0.2, linewidth = 0.8, na.rm = TRUE,
               position = position_dodge(width = 0.3)) +
  geom_jitter(aes(shape = Below_Detection), alpha = 0.5, 
              width = 0.15, height = 0, size = 2) +
  facet_grid(Measurement ~ Sample_Type, scales = "free_y", switch = "y") +
  scale_color_manual(values = color_palette_chicken,
                     name = "Biological\nReplicate") +
  scale_shape_manual(values = c("FALSE" = 16, "TRUE" = 4),
                     labels = c("Detected", "Non-detect"),
                     name = "Detection\nStatus") +
  labs(
    title = "Filtered Samples: Chicken Host-DNA Depletion Efficacy",
    subtitle = "Each point = qPCR technical replicate | × = Non-detect | Lines connect mean ± SE",
    x = NULL,
    y = NULL
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 10),
    strip.background = element_rect(fill = "gray90"),
    strip.text = element_text(face = "bold", size = 9),
    strip.placement = "outside",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
    legend.position = "right"
  )

print(chicken_panel)
ggsave("chicken_filtered_combined_panels.png", chicken_panel,
       width = 12, height = 10, dpi = 600)
```

## Statistical Analysis

```{r chicken-stats}
cat("\n========================================\n")
cat("CHICKEN DNA - STATISTICAL ANALYSIS\n")
cat("========================================\n\n")

# Function to analyze each sample type
analyze_filtered <- function(data, sample_type_name) {
  cat("\n--- ", sample_type_name, " ---\n\n", sep = "")
  
  paired_data <- data %>%
    filter(Sample_Type == sample_type_name) %>%
    select(Replicate_ID, Treatment, Mean_Ct, Mean_Conc, All_Non_Detect) %>%
    pivot_wider(names_from = Treatment,
                values_from = c(Mean_Ct, Mean_Conc, All_Non_Detect))
  
  before_conc <- paired_data$`Mean_Conc_Before Depletion`
  after_conc <- paired_data$`Mean_Conc_After Depletion`
  
  complete_pairs <- !is.na(before_conc) & !is.na(after_conc)
  n_complete <- sum(complete_pairs)
  
  if (n_complete >= 2) {
    before_complete <- before_conc[complete_pairs]
    after_complete <- after_conc[complete_pairs]
    
    # Choose test
    if (all(after_complete == 0) || all(after_complete < 0.0001)) {
      test_result <- wilcox.test(before_complete, after_complete,
                                 paired = TRUE, alternative = "greater")
      test_type <- "Wilcoxon signed-rank"
    } else {
      test_result <- t.test(before_complete, after_complete,
                           paired = TRUE, alternative = "greater")
      test_type <- "Paired t-test"
    }
    
    cat("  Test:", test_type, "\n")
    cat("  n pairs:", n_complete, "\n")
    cat("  Mean Before:", round(mean(before_complete), 4), "ng/μL\n")
    cat("  Mean After:", round(mean(after_complete), 6), "ng/μL\n")
    cat("  p-value:", format.pval(test_result$p.value, digits = 4), "\n")
    
    if (test_result$p.value < 0.001) {
      cat("  *** HIGHLY SIGNIFICANT ***\n")
    } else if (test_result$p.value < 0.01) {
      cat("  ** VERY SIGNIFICANT **\n")
    } else if (test_result$p.value < 0.05) {
      cat("  * SIGNIFICANT *\n")
    } else {
      cat("  Not significant\n")
    }
    
    # Effect size
    if (mean(after_complete) == 0 || mean(after_complete) < 0.0001) {
      cat("\n  Fold reduction: Complete (>10,000×)\n")
      cat("  Percent reduction: ~100%\n\n")
    } else {
      fold_change <- mean(before_complete) / mean(after_complete)
      percent_red <- ((mean(before_complete) - mean(after_complete)) / 
                      mean(before_complete)) * 100
      cat("\n  Fold reduction:", round(fold_change, 2), "×\n")
      cat("  Percent reduction:", round(percent_red, 1), "%\n\n")
    }
  } else {
    cat("  Insufficient pairs for testing\n\n")
  }
}

# Analyze both sample types
analyze_filtered(chicken_summary, "Centrifuge + Filter")
analyze_filtered(chicken_summary, "Filter Only")
```

---

# Campylobacter DNA Analysis

## Import and Clean Data

```{r campy-data}
# Read Campylobacter data
campy_raw <- read_excel('/Volumes/Alice/1_METHOD_DEVELOPMENT/Host-DNA_depletion/Filtered_samples_qPCR/qPCR_Filtered_Campylobacter.xlsx')

# Clean data
campy_clean <- campy_raw %>%
  filter(!is.na(Name), !str_detect(Name, "Blank")) %>%
  mutate(
    Treatment = if_else(str_detect(Name, "No Host"), "Before Depletion", "After Depletion"),
    Treatment = factor(Treatment, levels = c("Before Depletion", "After Depletion")),
    Sample_Type = case_when(
      str_detect(Name, "Cent \\+ Filt") ~ "Centrifuge + Filter",
      str_detect(Name, "^Filt") ~ "Filter Only",
      TRUE ~ "Unknown"
    ),
    Replicate_ID = as.numeric(str_extract(Name, "\\d+")),
    Below_Detection = is.na(Ct) | is.na(`Calc Conc (ng/ul)`),
    Ct_plot = ifelse(is.na(Ct), detection_limit_ct, Ct),
    Conc_plot = ifelse(is.na(`Calc Conc (ng/ul)`), detection_limit_conc, `Calc Conc (ng/ul)`)
  ) %>%
  select(Name, Treatment, Sample_Type, Replicate_ID, Ct, Ct_plot, 
         Concentration = `Calc Conc (ng/ul)`, Conc_plot, Below_Detection)

# Summary statistics
campy_summary <- campy_clean %>%
  group_by(Sample_Type, Treatment, Replicate_ID) %>%
  summarise(
    n_measurements = n(),
    n_detected = sum(!Below_Detection),
    n_non_detect = sum(Below_Detection),
    All_Non_Detect = all(Below_Detection),
    Mean_Ct = ifelse(all(is.na(Ct)), NA_real_, mean(Ct, na.rm = TRUE)),
    SE_Ct = ifelse(all(is.na(Ct)), NA_real_, sd(Ct, na.rm = TRUE) / sqrt(sum(!is.na(Ct)))),
    Mean_Conc = ifelse(all(is.na(Concentration)), NA_real_, mean(Concentration, na.rm = TRUE)),
    SE_Conc = ifelse(all(is.na(Concentration)), NA_real_, sd(Concentration, na.rm = TRUE) / sqrt(sum(!is.na(Concentration)))),
    .groups = "drop"
  )

campy_summary %>%
  select(Sample_Type, Treatment, Replicate_ID, n_detected, n_non_detect, Mean_Ct, Mean_Conc) %>%
  knitr::kable(digits = 4, caption = "Campylobacter DNA: Summary by Sample Type and Replicate")

campy_clean
```

## Combined Panel Plot

```{r campy-plot}
# Create long format
campy_long <- campy_clean %>%
  select(Sample_Type, Treatment, Replicate_ID, Below_Detection, Ct_plot, Conc_plot) %>%
  pivot_longer(
    cols = c(Ct_plot, Conc_plot),
    names_to = "Measurement",
    values_to = "Value"
  ) %>%
  mutate(
    Measurement = factor(Measurement,
                        levels = c("Ct_plot", "Conc_plot"),
                        labels = c("Ct Value", 
                                  "DNA Concentration (ng/µL)"))
  )

# Color palette
color_palette_campy <- c("1" = "#88D498", "2" = "#A846A0", "3" = "#98C1D9")

# Create plot
campy_panel <- ggplot(campy_long,
                      aes(x = Treatment, y = Value,
                          color = as.factor(Replicate_ID),
                          group = as.factor(Replicate_ID))) +
  stat_summary(fun = mean, geom = "line", linewidth = 1.2, na.rm = TRUE,
               position = position_dodge(width = 0.3)) +
  stat_summary(fun = mean, geom = "point", size = 3.5, na.rm = TRUE,
               position = position_dodge(width = 0.3)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", 
               width = 0.2, linewidth = 0.8, na.rm = TRUE,
               position = position_dodge(width = 0.3)) +
  geom_jitter(aes(shape = Below_Detection), alpha = 0.5, 
              width = 0.15, height = 0, size = 2) +
  facet_grid(Measurement ~ Sample_Type, scales = "free_y", switch = "y") +
  scale_color_manual(values = color_palette_campy,
                     name = "Biological\nReplicate") +
  scale_shape_manual(values = c("FALSE" = 16, "TRUE" = 4),
                     labels = c("Detected", "Non-detect"),
                     name = "Detection\nStatus") +
  labs(
    title = expression(paste("Filtered Samples: Host-DNA Depletion Effect on ", italic("Campylobacter"))),
    subtitle = "Each point = qPCR technical replicate | × = Non-detect | Lines connect mean ± SE",
    x = NULL,
    y = NULL
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 10),
    strip.background = element_rect(fill = "gray90"),
    strip.text = element_text(face = "bold", size = 9),
    strip.placement = "outside",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
    legend.position = "right"
  )

print(campy_panel)
ggsave("campylobacter_filtered_combined_panels.png", campy_panel,
       width = 12, height = 10, dpi = 600)
```

## Statistical Analysis

```{r campy-stats}
cat("\n========================================\n")
cat("CAMPYLOBACTER DNA - STATISTICAL ANALYSIS\n")
cat("========================================\n\n")

analyze_filtered(campy_summary, "Centrifuge + Filter")
analyze_filtered(campy_summary, "Filter Only")
```

---

# Salmonella DNA Analysis

## Import and Clean Data

```{r salm-data}
# Read Salmonella data
salm_raw <- read_excel('/Volumes/Alice/1_METHOD_DEVELOPMENT/Host-DNA_depletion/Filtered_samples_qPCR/qPCR_filtered-host-dep_Salmonella.xlsx')

# Clean data
salm_clean <- salm_raw %>%
  filter(!is.na(Name), !str_detect(Name, "Blank")) %>%
  mutate(
    Treatment = if_else(str_detect(Name, "No Host"), "Before Depletion", "After Depletion"),
    Treatment = factor(Treatment, levels = c("Before Depletion", "After Depletion")),
    Sample_Type = case_when(
      str_detect(Name, "Cent \\+ Filt") ~ "Centrifuge + Filter",
      str_detect(Name, "^Filt") ~ "Filter Only",
      TRUE ~ "Unknown"
    ),
    Replicate_ID = as.numeric(str_extract(Name, "\\d+")),
    Below_Detection = is.na(Ct) | is.na(`Calc Conc (ng/ul)`),
    Ct_plot = ifelse(is.na(Ct), detection_limit_ct, Ct),
    Conc_plot = ifelse(is.na(`Calc Conc (ng/ul)`), detection_limit_conc, `Calc Conc (ng/ul)`)
  ) %>%
  select(Name, Treatment, Sample_Type, Replicate_ID, Ct, Ct_plot, 
         Concentration = `Calc Conc (ng/ul)`, Conc_plot, Below_Detection)

# Summary statistics
salm_summary <- salm_clean %>%
  group_by(Sample_Type, Treatment, Replicate_ID) %>%
  summarise(
    n_measurements = n(),
    n_detected = sum(!Below_Detection),
    n_non_detect = sum(Below_Detection),
    All_Non_Detect = all(Below_Detection),
    Mean_Ct = ifelse(all(is.na(Ct)), NA_real_, mean(Ct, na.rm = TRUE)),
    SE_Ct = ifelse(all(is.na(Ct)), NA_real_, sd(Ct, na.rm = TRUE) / sqrt(sum(!is.na(Ct)))),
    Mean_Conc = ifelse(all(is.na(Concentration)), NA_real_, mean(Concentration, na.rm = TRUE)),
    SE_Conc = ifelse(all(is.na(Concentration)), NA_real_, sd(Concentration, na.rm = TRUE) / sqrt(sum(!is.na(Concentration)))),
    .groups = "drop"
  )

salm_summary %>%
  select(Sample_Type, Treatment, Replicate_ID, n_detected, n_non_detect, Mean_Ct, Mean_Conc) %>%
  knitr::kable(digits = 4, caption = "Salmonella DNA: Summary by Sample Type and Replicate")
```

## Combined Panel Plot

```{r salm-plot}
# Create long format
salm_long <- salm_clean %>%
  select(Sample_Type, Treatment, Replicate_ID, Below_Detection, Ct_plot, Conc_plot) %>%
  pivot_longer(
    cols = c(Ct_plot, Conc_plot),
    names_to = "Measurement",
    values_to = "Value"
  ) %>%
  mutate(
    Measurement = factor(Measurement,
                        levels = c("Ct_plot", "Conc_plot"),
                        labels = c("Ct Value", 
                                  "DNA Concentration (ng/µL)"))
  )

salm_long

# Color palette
color_palette_salm <- c("1" = "#88D498", "2" = "#A846A0", "3" = "#98C1D9")

# Create plot
salm_panel <- ggplot(salm_long,
                     aes(x = Treatment, y = Value,
                         color = as.factor(Replicate_ID),
                         group = as.factor(Replicate_ID))) +
  stat_summary(fun = mean, geom = "line", linewidth = 1.2, na.rm = TRUE,
               position = position_dodge(width = 0.3)) +
  stat_summary(fun = mean, geom = "point", size = 3.5, na.rm = TRUE,
               position = position_dodge(width = 0.3)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", 
               width = 0.2, linewidth = 0.8, na.rm = TRUE,
               position = position_dodge(width = 0.3)) +
  geom_jitter(aes(shape = Below_Detection), alpha = 0.5, 
              width = 0.15, height = 0, size = 2) +
  facet_grid(Measurement ~ Sample_Type, scales = "free_y", switch = "y") +
  scale_color_manual(values = color_palette_salm,
                     name = "Biological\nReplicate") +
  scale_shape_manual(values = c("FALSE" = 16, "TRUE" = 4),
                     labels = c("Detected", "Non-detect"),
                     name = "Detection\nStatus") +
  labs(
    title = expression(paste("Filtered Samples: Host-DNA Depletion Effect on ", italic("Salmonella"))),
    subtitle = "Each point = qPCR technical replicate | × = Non-detect | Lines connect mean ± SE",
    x = NULL,
    y = NULL
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 10),
    strip.background = element_rect(fill = "gray90"),
    strip.text = element_text(face = "bold", size = 9),
    strip.placement = "outside",
    axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
    legend.position = "right"
  )

print(salm_panel)
ggsave("salmonella_filtered_combined_panels.png", salm_panel,
       width = 12, height = 10, dpi = 600)
```

## Statistical Analysis

```{r salm-stats}
cat("\n========================================\n")
cat("SALMONELLA DNA - STATISTICAL ANALYSIS\n")
cat("========================================\n\n")

analyze_filtered(salm_summary, "Centrifuge + Filter")
analyze_filtered(salm_summary, "Filter Only")
```

---

# Summary

```{r summary}
cat("\n========================================\n")
cat("FINAL SUMMARY\n")
cat("========================================\n\n")

cat("Three high-resolution plots generated:\n")
cat("1. chicken_filtered_combined_panels.png\n")
cat("2. campylobacter_filtered_combined_panels.png\n")
cat("3. salmonella_filtered_combined_panels.png\n\n")

cat("Key findings:\n")
cat("- Chicken DNA (host): Expected to be REDUCED after depletion\n")
cat("- Campylobacter DNA (pathogen): Expected to be RETAINED/INCREASED\n")
cat("- Salmonella DNA (pathogen): Expected to be RETAINED/INCREASED\n\n")

cat("Statistical tests automatically selected:\n")
cat("- Wilcoxon test when after values near zero\n")
cat("- Paired t-test otherwise\n\n")

sessionInfo()
```
